---
title: "HW_5_Grebeniuk"
author: "Grebeniuk_Alena"
date: "12/2/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
# Exercise 1 . Pre-process the data so that you have the following four columns:
## subj compatible correct rt
```{r}

getwd()
setwd("/Users/alenagrebenuk/Desktop/stats1_hw")
dat<-read.table("stroopdata.txt",header=FALSE)
head(dat)
dat<-dat[,c(3,4,7,8)]
head(dat)
colnames(dat)<-c("subj","compatible","correct","rt")
head(dat)
```

## subj refers to subject id (factor) 
```{r}
dat$subj<-as.factor(dat$subj)
```

## compatible is coded as 1 (compatible) or 0 (incompatible), also a factor
```{r}
dat$compatible<-as.factor(dat$compatible)
```

## correct (whether the response is correct or incorrect), this is an integer (numeric)
```{r}
dat$correct<-as.integer(dat$correct)
```

## rt is the reaction time in milliseconds (also an integer, numeric)
```{r}
dat$rt<-as.integer(dat$rt)
```

# Ecercise 2 
## Plot the distribution of reaction times for the compatible and incompatible conditions using histograms and boxplots. 

```{r}
compatible_rt<-subset(dat,compatible==1)$rt
incompatible_rt<-subset(dat, compatible==0)$rt
hist(compatible_rt)
hist(incompatible_rt)
```
```{r}
boxplot(compatible_rt)
boxplot(incompatible_rt)

```
## the thick horizontal line in the box is MEDIAN - (the middle value of the dataset).
## what are the ends of the whiskers? -THOSE ARE ENDS OF OBSERVED DATA POINTS - (Minimum and maximum values within a certain range).
## and what do the top and bottom edges of the box represent? - the Bottom is the Rirst Quartile (Q1)(he median of the lower half of the data) 
## the Top Edge is the Third Quartile (Q3)(the median of the upper half of the data.)


# Exercise 3 
## What is the mean difference between the incongruent and incongruent conditions (across all subjects)?
```{r}
mean(subset(dat,compatible==1)$rt)
```


```{r}

mean(subset(dat,compatible==0)$rt)
```

# Exercise 4 
## For each subject, what are the means differences between the incongruent and incongruent conditions?
```{r}

with(dat,tapply(rt, IND=list(subj,compatible),mean))

getwd()
```

# Exercise 5 
## What do you conclude from the differences in means? Can we conclude that incongruent conditions have longer reaction times than congruent conditions?

## we can coclude that incongruent conditions have longer reaction times than congruent conditions. Since the difference in mean showed that 1108.1833 > 982.6296. 

_heello - to make ithe text cursive 

> Hvjbkjhbk - to make the "chunk" for the text

```{r}


l1 <- c(1, 2, 3, 6, 10)
l2 <- c(6, 2, 4, 5, 55)

print(l1+l2)
```

```{r}
getwd()
SE<-31.826/sqrt(42)
print(SE)

```
```{r}
# Given values
sample_mean <- 12.779
hypothesized_mean <- 0
sample_sd <- 31.826
sample_size <- 42

# Calculate the t-value
observed_t_value <- abs((sample_mean - hypothesized_mean) / (sample_sd / sqrt(sample_size)))

print(observed_t_value)

# Print the result
```





```{r}
Sd<- (24.513-0)/4.668
Sd
SE<-5.251/sqrt(15)
SE
```





```{r}
t_value<-2.602
df=41
p_value<-2*pt(abs(t_value), df, lower.tail=FALSE)
print(p_value)
```




```{r}
# Given sample size
n <- 42

# Calculate degrees of freedom
df <- n - 1

# Print the result
print(df)

```
```{r}
# Given significance level (e.g., 0.05 for a 95% confidence level)
alpha <- 0.05

# Given degrees of freedom (df)
df <- 41  # replace with your actual degrees of freedom

# Calculate critical t-value
critical_t <- qt(1 - alpha/2, df)

# Take the absolute value
absolute_critical_t <- abs(critical_t)

# Print the result
print(absolute_critical_t)

```

```{r}
# Example code (replace df and crit_val with your values)
df <- 30  # degrees of freedom
crit_val <- 2.042  # critical t-value

# Probability of observing a value as extreme as or more extreme than crit_val
p_value <- pt(crit_val, df, lower.tail = FALSE)

# Print the result
print(p_value)

```
```{r}
# Example code (replace obs_t, df with your values)
obs_t <- 2.5  # observed t-value
df <- 30  # degrees of freedom

# Probability of observing a t-value as extreme as or more extreme than obs_t
p_value_two_tailed <- 2 * (1 - pt(abs(obs_t), df))

# Print the result
print(p_value_two_tailed)

```
```{r}
# Assuming df is the degrees of freedom
df <- n-1  # Replace ... with the actual degrees of freedom

# Calculate critical t-value
critical_t_value <- qt(0.025, df, lower.tail = FALSE)
print(critical_t_value )
```

```{r}
# Given values
sample_mean <- 24.513
abs_observed_t_value <- 4.668
sample_size <- 15
mu<- 0

# Degrees of freedom 0
df <- sample_size - 1

# Critical t-value for a two-sided test at a 0.05 significance level
critical_t_value <- abs(qt(0.025, df))
#absolute_critical_t <- abs(critical_t_value)
critical_t_value
# Print the result
#print(absolute_critical_t)

# Estimated standard error
SE <- abs_observed_t_value / sample_mean
SE
# Estimated standard deviation
SD <- SE * sqrt(sample_size)

# P-value for a two-sided t-test
p_value <- 2 * pt(-abs(observed_t_value), df)


```

```{r}
# Given values
alpha <- 0.05
df <- 14
absolute_observed_t_value <- 4.668

# Calculate the critical t-value for a two-sided test
absolute_critical_t_value <- qt(1 - alpha / 2, df)

# Output the result
cat("Absolute Critical t-value:", absolute_critical_t_value, "\n")

```
```{r}
sample_mean <- 24.513
n<-15
SEE<-sd(sample_mean)/(sqrt(n))
SE<-90/sqrt(15)
SE
SEE
```

```{r}
observed_t<-2.019
n<-42
pt(observed_t,df=n-1,lower.tail=FALSE)
```


```{r}
# Given data
sample_mean <- 24.513
sample_size <- 15  # Replace with the actual sample size
confidence_level <- 0.95
alpha_over_2 <- (1 - confidence_level) / 2
degrees_of_freedom <- 14  # Replace ... with the actual degrees of freedom

# Critical t-value
critical_t <- qt(1 - alpha_over_2, df = degrees_of_freedom)

# Calculate estimated standard error
estimated_se <- sample_mean / (critical_t / sqrt(sample_size))

# Print the result
print(estimated_se)

```

```{r}
# Assuming you already have estimated_se and sample_size
estimated_sd <- estimated_se * sqrt(sample_size)

# Print the result
print(estimated_sd)

```

```{r}
# Assuming you already have abs_observed_t and df
#p_value = 2 * (1 - pt(abs_observed_t, df))
n<-15
p_value<-2*pt(abs(observed_t),df=n-1,
              lower.tail=FALSE)
# Print the result
print(p_value)

```

```{r}
# Example values
sample_mean <- 24.513
population_mean <- 0  # Assuming null hypothesis is that population mean is 0
sample_sd <- 171.436 # Replace with your actual sample standard deviation
sample_size <- 15

# Calculate standard error
standard_error <- sample_sd / sqrt(sample_size)

# Calculate observed t-value
observed_t <- (sample_mean - population_mean) / standard_error

# Print the result
print(observed_t)

```
```{r}
 dat<-read.table("stroopdata.txt",header=FALSE)

 dat$subj<-rep(1:40,each=40)

 dat<-dat[,c(4,7,8,9)]

 colnames(dat)<-c("compatible","response","rt","subj")

 dat$subj<-factor(dat$subj)

 dat$compatible<-factor(dat$compatible)

 dat<-subset(dat,response!=3)

 dat$correct<-ifelse(dat$response==1,1,0)

 means<-with(dat,tapply(rt,IND=list(subj,compatible),mean))

 head(means)
```

```{r}
diff<-means[,2]-means[,1]
```


```{r}
round(mean(diff),3)
```

```{r}

print(sd)
SE<-
abs_t_valaue<-
abs_critical_t_value<-
p_value_two_sided<-
  
# Given values
sample_mean <- 24.513
observed_t_value <- 4.668
sample_size <- 15

# Degrees of freedom 
df <- sample_size - 1

# Critical t-value for a two-sided test at a 0.05 significance level
critical_t_value <- qt(0.025, df)

# Estimated standard error
SE <- observed_t_value / sample_mean

# Estimated standard deviation
SD <- SE * sqrt(sample_size)

# P-value for a two-sided t-test
p_value <- 2 * pt(-abs(observed_t_value), df)
p_value
```


```{r}
SE<-14.254
sample_mean<- -125.472
abs_t_valaue<-abs(sample_mean/SE)
print(abs_t_valaue)
```


```{r}
abs_t_valaue<-8.802
n<-40
p_value<-2*pt(abs_t_valaue, df=n-1,lower.tail = FALSE)
p_value
```



```{r}
# Example data


# Hypothetical mean under the null hypothesis
null_mean <- 0

# Perform a t-test
t_test_result <- t.test( rt ~ dat$compqtible, mu = null_mean)

# Extract the observed t-value
observed_t_value <- t_test_result$statistic

# Output the result
cat("Observed t-value:", observed_t_value, "\n")

```



```{r}
# Calculate the standard deviation of differences
sd_diff <- sd(diff)
print(sd_diff)
# Calculate the square root of the sample size
sqrt_n <- sqrt(length(diff))
print(sqrt_n)
# Calculate the standard error
standard_error <- sd_diff / sqrt_n
print(standard_error)


```
```{r}
# Assuming you have 'dat' and 'means' as defined in the previous R code

# Calculating the observed t-value
t_result <- t.test(rt ~ compatible, data = dat)

# Extracting the observed t-value
observed_t_value <- t_result$statistic

# Display the observed t-value
observed_t_value

```
```{r}
# Assuming you have 'observed_t_value' as calculated before

# Calculating the absolute observed t-value
absolute_observed_t_value <- abs(observed_t_value)

# Display the absolute observed t-value
absolute_observed_t_value

```

```{r}
n<-40
abs_critical_t<-abs(qt(0.025,df=n-1))
abs_critical_t
```

```{r}
2*pt(abs(observed_t),df=n-1,lower.tail=FALSE)
```


```{r}
df<-41
aaa<-pt(2.019, df)
aaa

```
```{r}
# Given values
absolute_observed_t_value <- 4.668
sample_mean <- 24.513
sample_size <- 15

# Calculate the estimated standard error
estimated_standard_error <- absolute_observed_t_value / sample_mean

# Calculate the estimated standard deviation
estimated_standard_deviation <- estimated_standard_error * sqrt(sample_size)

# Output the results
cat("Estimated Standard Error:", estimated_standard_error, "\n")
cat("Estimated Standard Deviation:", estimated_standard_deviation, "\n")

```
```{r}
sd(24.513)
```

```{r}
observed_t<-8.802
n<-40
2*pt((observed_t),df=n-1,lower.tail=FALSE)
```


```{r}
observed_t<-8.802
n<-40
hello<-pt(observed_t,df=n-1,lower.tail = FALSE)
hello
```


```{r}
abs_observed_t<-2.022
n<-40
pt(abs_observed_t,df=n-1, lower.tail=FALSE)+pt(-abs_observed_t,df=n-1)
```

```{r}
n<-15
sample_mean<-sd(n)
SE<-sd(sample_mean)/sqrt(n)
sample_mean
SE
```

```{r}
diff_sim<-rnorm(n=15,mean=24.513,sd=90) 
stddev<-sd(diff_sim)
SE<-stddev/sqrt(15)
diff_sim
stddev
SE
```

